\plaintitle

### 2016-07-28 drafts/labor/introduction.tex
major revisions based on flash meeting feedback

% %!TEX root = ./trackingWork.tex
\documentclass[trackingWork]{subfiles}
\makeatletter
\def\blx@maxline{77}
\makeatother
\begin{document}

\newcommand{\note}[1]{\onlyinsubfile{\textit{#1}}}

\section{Introduction}
% each sentence feels like a thought that's unrelated to the next one
The past decade has seen a proliferation in on--demand work
where workers' statuses are so fleeting that
they have been described as ``transient''.
After identifying that myriad tasks
--- from the mundane to the complex ---
can be accomplished
% beginning with ``information work'' such as on
% Amazon Mechanical Turk (AMT).
% ??? ->
% Stemming from the recognition that
% complex tasks might be achievable
% --- and indeed trivial ---
by directing and managing the ``crowd'',
the research and and industry communities have flocked to sites of labor
like Amazon's Mechanical Turk (AMT)
to solicit workers.
Researchers in particular have taken to the space in earnest,
finding opportunities to enable new forms of work
as well as using these workers (``Turkers'')
as representative populations of the public
\cite{bernsteinSoylent,redballoon,paolacci2010running}.

\note{Is this different from other contract work? yes. here's why:}

This approach to managing work (and in particular workers)
can be differentiated from other formulations by a constellation of traits:
\begin{inlinelist}
\item distributing the workforce, % gray
\item distancing workers from the context of the work, and
\item compensating work on a per--task basis rather than for time
% citations! future of crowdwork or mary's or lily's work / quin and bederson
\end{inlinelist}.
These features can be found elsewhere
--- for instance,
distributed workforces
are a feature of global labor markets,
and abstracting workers from the products of their labor
is well--associated with
\citeauthor{marx2012economic} ---
but these components have been explored only individually
(for instance, by \citeauthor{fevre1986contract} and later \citeauthor{kalleberg2000nonstandard}),
and we focus on the newfound interest in
work arranging these components in what holistically seems to be a new mode of work
% While contract work is familiar enough,
% much of this work focuses on comprehensive work
% (for example, creative or professional work like landscaping, design, etc\dots)
\cite{kalleberg2000nonstandard,fevre1986contract,marx2012economic}.

\note{why is this important? because it's bigger than AMT}

On--demand work under this formulation
has sparked interest across industries ranging from
livery (driving for hire, for example Uber),
house--cleaning (Handy),
and various other services (TaskRabbit)
\cite{uberOfficial,zaarlyOfficial,taskrabbitOfficial}.
Today, a rapidly growing transient workforce is forming,
itself assembling piece--by--piece as industries and researchers find more and more unexpected ways
to benefit from a latent pool of variably skilled workers.

\note{this is getting into more general ``how has this moved forward'' kind of work. save for later?}

Broadly, this work has imagined the various applications of ``gig''--based work,
and making it work for ``requesters''
--- academics, industries, and to some extent end--users
who might benefit from the product of crowd--powered work
\cite{hong2015group,jonBrelig,paolacci2010running}.
% ??? -->
Since then,
the application of digitally mediated environments acting as rapid ``hiring halls''
for on--demand tasks
has found purchase far beyond the context of ``information work'' like Amazon Mechanical Turk (AMT);
livery services (Uber), cleaning (Handy), and various other services (TaskRabbit)
% parentheticals giving examples?
have garnered the interest of both industry and research.

Researchers have made efforts to understand the \textit{people} that have gravitated toward on--demand work
since its emergence and popularization,
but as the form of work has grown and changed, so too have the demographics of workers
\cite{Ross,whoareNOTtheTurkers}.
Some of this research has been motivated by the realization of the sociality of gig work,
and the frustration and disenfranchisement that these systems embody
\cite{turkopticon,dynamo}.
Other work has focused on the outcomes of work,
reflecting on the resistance workers express against digitally mediated labor markets
\cite{uberAlgorithm}.

Bringing this broader body of work together, perhaps we can think of this as documenting the ``fallout'' of on--demand labor;
workers increasingly expressing frustration with the disenfranchisement of the work--flows pioneered by researchers in the decade leading up till now,
and researchers beginning to appreciate the unforeseen costs of transient labor and the impacts both on the work and the workers.

\note{this probably doesn't belong here but i need to put it somewhere!}

From various perspectives and with an eye toward myriad stakeholders,
this body of work has ostensibly sought to answer one overarching question:
What does the future hold for work and those that do it? % expand on ...?
% make it clearer
% inevitable/predictable outcomes?
% aspirational -> how can we make work better for people
Whether that research has attempted to articulate the potential to achieve difficult and creative tasks,
or wrestled with the frustration that workers face on a near--constant basis,
these roads of inquiry have, it seems, looked to the horizon to try to predict where we are heading.

Rather than looking necessarily to the creation of systems
(both as genuine agents and as provocateurs (e.g. Soylent and Turkopticon)),
we turn to the \textit{past}, uncovering ``piecework'' as
a useful model for understanding the mode of work that has captured the imagination of entrepreneurs and researchers.

\subsection{Piecework as a lens to understand gig work}
These bodies of literature have conversed,
and indeed begun to identify that their theories are interrelated in some way,
but thus far a convincing model for making sense of
--- let alone predicting ---
the phenomena in on--demand labor has proven elusive.
For instance, 
making sense of why workers on AMT (or ``Turkers'') are frustrated with their workplace
is understandable in its own context, but
the relationship between the work itself (``turking'')
and the myriad social outcomes has only briefly been explored
\cite{crowdcollab}.

We offer a framing for \textit{all} on--demand work as a modern instantiation of ``piecework'',
referenced briefly by
\citeauthor{crowdworkFuture}
in
\citeyear{crowdworkFuture}
as a loose analogy to the form of work emerging at the time,
but not given as much weight as it perhaps could have been
\cite{crowdworkFuture}.
``Piecework'' here makes use of
a term historically used to describe work done in the home,
in manageable tasks,
often involving clear instructions
and payment only for work completed, not work done
(the differentiation then being that
one would be paid for the \textit{output} of the work,
not the \textit{duration}).

This framing prompts us to note several immediately apparent similarities:
% Given the scope,
% we can frame piecework and on--demand labor
% as sharing these important similarities:
\begin{inlinelist}
\item this form of work began in the home
\item the worker is paid for each discrete piece of work done, regardless of time or effort; and
\item the worker's status
(not only socially, but also economically)
is ambiguous, or at least the subject of some controversy.
\end{inlinelist}

% In the past decade, researchers have observed frustration
% grow among on--demand workers,
% with expression of this frustration spanning a wide range of tactics
% \cite{uberAlgorithm,turkopticon,dynamo}.
% Attempting to make sense of these case studies has been challenging
% in part because
% a wholly encompassing framework for understanding this form of work
% has thus proven difficult to capture.

>>>>

\documentclass[trackingWork]{subfiles}
\makeatletter
\def\blx@maxline{77}
\makeatother
\begin{document}

\newcommand{\note}[1]{\onlyinsubfile{\ali{#1}}}
\section{Introduction}
The past decade has seen a flourishing of on--demand work
where the statuses of workers have become so fleeting that 
these workers
(known colloquially as ``Turkers'')
have been described as ``transient''
\cite{kargerIterativeLearning,mitraComparingStrategies,latozaCrowdDev}.
The realization
that tasks can be accomplished by directing and managing this crowd of workers
has spurred the research and and industry communities to flock to sites of labor
like Amazon's Mechanical Turk (AMT)
to explore the limits of this distributed, seemingly ephemeral labor force.
Researchers in particular have taken to the space in earnest,
finding opportunities to enable new forms of work
as well as using Turkers
as representative populations of the public
\cite{bernsteinSoylent,redballoon,paolacci2010running}.

The many sites of work replicating and extending on
the general style of labor popularized by AMT
have predominantly involved work done on a computer
or involving the human processing of data,
leading many to call this ``information work"
\cite{turkopticon,professionalCrowdworkEthics,IraniFromCriticalDesign,OlsonMakeDistanceWorkWork}.
\citeauthor{howe2008crowdsourcing} defined ``crowdsourcing'' in \citeyear{howe2008crowdsourcing}
as
``taking a job traditionally performed by a designated agent
(usually an employee)
and outsourcing it to an undefined,
generally large group of people in the form of an open call''
\cite{howe2008crowdsourcing}.

In the years since,
scholars have generated taxonomies for the work done by many distributed workers
in an attempt to better categorize and reason about the many forms of work done 
on information work platforms such as AMT, oDesk, etc\dots
\cite{yuenSurvey,geiger2011managing,quinnbedersonTaxonomy}.
We add that,
under \citeauthor{howe2008crowdsourcing}'s constraints,
even \textit{more} new forms of work fall squarely under
the metaphorical umbrella we collectively call ``crowdsourcing''.

Indeed, this on--demand workforce has sparked interest
across industries ranging
livery (driving for hire, for example Uber),
house--cleaning (Handy),
and generalized services (TaskRabbit)
\cite{uberOfficial,zaarlyOfficial,taskrabbitOfficial}.
Today, a rapidly growing transient workforce is forming,
itself assembling piece--by--piece as industries and researchers find yet more unexpected ways
to benefit from a latent pool of previously vetted workers
\cite{pewSharing}.

% Broadly, this work has imagined the various applications of ``gig''--based work,
% and making it work for ``requesters''
% --- academics, industries, and to some extent end--users
% who might benefit from the product of crowd--powered work
% \cite{hong2015group,jonBrelig,paolacci2010running}.
% % ??? -->
% Since then,
% the application of digitally mediated environments acting as rapid ``hiring halls''
% for on--demand tasks
% has found purchase far beyond the context of ``information work'' like Amazon Mechanical Turk (AMT);
% livery services (Uber), cleaning (Handy), and various other services (TaskRabbit)
% % parentheticals giving examples?
% have garnered the interest of both industry and research.

Researchers have made efforts to understand the people
that have gravitated toward crowdsourcing platforms
since its emergence and popularization,
but as the form of work has grown and changed, so too have the demographics of workers
\cite{Ross,whoareNOTtheTurkers}.
Some of this research has been motivated by the realization of the sociality of gig work,
and the frustration and disenfranchisement that these systems embody
\cite{turkopticon,dynamo}.
Other work has focused on the outcomes of work,
reflecting on the resistance workers express against digitally mediated labor markets
\cite{uberAlgorithm}.

% Bringing this broader body of work together, perhaps we can think of this as documenting the ``fallout'' of on--demand labor;
% workers increasingly expressing frustration with the disenfranchisement of the work--flows pioneered by researchers in the decade leading up till now,
% and researchers beginning to appreciate the unforeseen costs of transient labor and the impacts both on the work and the workers.

% \note{this probably doesn't belong here but i need to put it somewhere!}

From various perspectives and with an eye toward myriad stakeholders,
this body of work has ostensibly sought to answer one overarching question:
What does the future hold for work and those that do it? % expand on ...?
To give several key examples,
the driving motivation behind much of the research in this space has led us to ask,
at a high level, three important questions:
\begin{enumerate}
  \item What are the (creative, cognitive, emotional) limits of crowdsourcing?
    \cite{foundry,suzukiAtelier,KimStoria,yuanAlmost,YuEncouragingOutside,embracingErrorKrishna,Nebeling:2016:WCW:2858036.2858169,Hahn:2016:KAB:2858036.2858364};
  \item What forms of worker management and arrangement are viable (and moreover \textit{effective})?
    \cite{bernsteinSoylent,sensitiveTasks,LykourentzouPersonalityMatters,KucherbaevReLauncher,Law:2016:CKC:2858036.2858144,Cai:2016:CRI:2858036.2858237,Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}; and%
  \item What will work and the place of work look like for the workers?
    \cite{turkopticon,storiesIraniSilberman,dynamo,crowdcollab,whyWouldAnyoneBrewer,takingAHITMcInnis}.
    \note{these questions are all a little guilty of vagueness but this feels especially ``meh''}
\end{enumerate}

\subsection{Piecework as a lens to understand crowdsourcing}
This large and growing body of research has conversed
to varying degrees with labor scholarship,
but has not offered a persuasive framing for holistically explaining
the developments in worker processes that researchers have developer, or
the phenomena in social environments we have observed;
nor has any research, to our knowledge,
gone as far as predict future developments.

% These bodies of literature have conversed,
% and indeed begun to identify that their theories are interrelated in some way,
% but thus far a convincing model for making sense of
% --- let alone predicting ---
% the phenomena in on--demand labor has proven elusive.
% For instance, 
% making sense of why workers on AMT (or ``Turkers'') are frustrated with their workplace
% is understandable in its own context, but
% the relationship between the work itself (``turking'')
% and the myriad social outcomes has only briefly been explored
% \cite{crowdcollab}.

We offer a framing for crowd work spanning the aforementioned industries
collectively as a contemporary instantiation of ``piecework''.
Piecework as a metaphor for the type of work at hand is not new
(indeed,
\citeauthor{crowdworkFuture}
in
\citeyear{crowdworkFuture}
referenced crowd work as ``piecework'' briefly
as a loose analogy to the form of work emerging at the time),

% but not given as much weight as it perhaps could have been
\cite{crowdworkFuture}.
% ``Piecework'' here makes use of
% a term historically used to describe work done in the home,
% in manageable tasks,
% often involving clear instructions
% and payment only for work completed, not work done
% (the differentiation then being that
% one would be paid for the \textit{output} of the work,
% not the \textit{duration}).

% This framing prompts us to note several immediately apparent similarities:
% \begin{inlinelist}
%   \item this form of work began in the home
%   \item the worker is paid for each discrete piece of work done, regardless of time or effort; and
%   \item the worker's status
%   (not only socially, but also economically)
%   is ambiguous, or at least the subject of some controversy.
% \end{inlinelist}

### 2016-07-28 drafts/labor/trackingWork.tex
more intro revisions
major revisions based on flash meeting feedback

@@

\begin{abstract}
  With growing attention toward on--demand labor
  --- ranging from the ``sharing economy'' to information work ---
  scholars have made connections to various frameworks 
  and mechanisms such as worker advocacy, empowerment, and Taylorism,
  to make sense of our observations of
  on--demand work and the workers that power this movement.
  We argue that the literature surrounding ``piecework''
  informs and even predicts both the contributions
  that have been made toward the development of on--demand labor and crowd work
  as well as the fallout among workers and researchers
  with regard to the disillusionment and alienation of work.

@@

Work decomposition, then, is far from new;
``decomposition'' generally illustrates the same concepts of work that ``Taylorism''
and scientific management sought to embody
---
\citeauthor{silberman2010ethics} in particular foresaw this danger and warned of it in
\citeyear{silberman2010ethics}
\cite{crowdworkFuture,silberman2010ethics,nickerson2013crowd}.
In both the historical and contemporary cases of decomposed work,
work was,
at least initially,
distributed in the form of tasks to the homes of workers;
\citeauthor{riisOtherSideLives} captured this in his documentary work
\citetitle{riisOtherSideLives}
in \citeyear{riisOtherSideLives}
\cite{riisOtherSideLives}.

@@

The scientific management of pieceworkers has been well--studied under the umbrella of assembly line research,
and even physiological study of pieceworkers closely resembles the
research into cognitive loads and stress levels that we discussed among on--demand crowd workers
\cite{pieceworkBiologicalHarm,hu1961parallel}.
Even the complicated relationships between observers and workers themselves are not necessarily new;
\citeauthor{riisOtherSideLives}'s photodocumentary of pieceworkers has even been re--examined through an exercise asking crowd workers
to photograph themselves for similar purposes as \citeauthor{riisOtherSideLives}'s --- to document and humanize an otherwise abstracted, invisible workforce
\cite{facesOfMechanicalTurk,turkopticon,riisOtherSideLives}.

>>>>

\setlist{nolistsep}

@@

\begin{abstract}
  With growing attention toward on--demand labor
  --- ranging from the ``sharing economy'' to information work ---
  scholars have made connections to various frameworks 
  and mechanisms such as worker advocacy, empowerment, and Taylorism,
  to make sense of our observations of
  on--demand work and the workers that power this movement.
  We argue that the literature surrounding ``piecework''
  informs and even predicts both the contributions \cite{}
  that have been made toward the development of on--demand labor and crowd work
  as well as the fallout among workers and researchers
  with regard to the disillusionment and alienation of work.

@@

Work decomposition, then, is far from new;
``decomposition'' generally illustrates the same concepts of work that ``Taylorism''
and scientific management sought to embody
---
\citeauthor{professionalCrowdworkEthics} in particular foresaw this danger and warned of it in
\citeyear{professionalCrowdworkEthics}
\cite{crowdworkFuture,professionalCrowdworkEthics,nickerson2013crowd}.
In both the historical and contemporary cases of decomposed work,
work was,
at least initially,
distributed in the form of tasks to the homes of workers;
\citeauthor{riisOtherSideLives} captured this in his documentary work
\citetitle{riisOtherSideLives}
in \citeyear{riisOtherSideLives}
\cite{riisOtherSideLives}.

@@

The scientific management of pieceworkers has been well--studied under the umbrella of assembly line research,
and even physiological study of pieceworkers closely resembles the
research into cognitive loads and stress levels that we discussed among on--demand crowd workers
\cite{hu1961parallel,pieceworkBiologicalHarm}.
Even the complicated relationships between observers and workers themselves are not necessarily new;
\citeauthor{riisOtherSideLives}'s photodocumentary of pieceworkers has even been re--examined through an exercise asking crowd workers
to photograph themselves for similar purposes as \citeauthor{riisOtherSideLives}'s --- to document and humanize an otherwise abstracted, invisible workforce
\cite{facesOfMechanicalTurk,turkopticon,riisOtherSideLives}.
